---
title: "Data Wrangling Pt 3: Time Series Data and Missing Values"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
---

## Load Packages 

```{r chunk01}
library(dplyr)

# Specify package preferences for filter, count, select, and arrange
library(conflicted)
conflict_prefer("filter", "dplyr", quiet = TRUE)
conflict_prefer("count", "dplyr", quiet = TRUE)
conflict_prefer("select", "dplyr", quiet = TRUE)
conflict_prefer("arrange", "dplyr", quiet = TRUE)
```

\

## Load some weather data from the CIMIS network

```{r chunk02}
library(googlesheets4)
arvinedison_url <- "https://docs.google.com/spreadsheets/d/13FR4Mji24KqZmLqK_BiLKew8nN4MbCdh1KpzT54hH4M/"

## Tell R not to authorize with Google (not needed for this public Google Sheet)
gs4_deauth()

arvinedison_tbl <- read_sheet(arvinedison_url, sheet = "Sheet1")

## View(arvinedison_tbl)
head(arvinedison_tbl)
```

**Pro Tip:**

You can download data from the CIMIS network directly into R using the [`cimir`](https://cran.r-project.org/package=cimir) package (Michael Koohafkan).

```
library(cimir)
set_key("xxxxxxxx-xxxx-xxxx-xxxx-xxxxxxxxxxxx")
arvinedison_tbl <- cimis_data(targets = 125, start.date = "2023-09-16", end.date = "2023-10-15",
                               items = "day-air-tmp-max,day-air-tmp-min,day-eto,day-precip")
```

\

## CHALLENGE #1. Would you consider the CIMIS data to be long or wide? It is tidy?

\

## Create a date column

```{r chunk03}
library(lubridate)
arvedsn_dt_tbl <- arvinedison_tbl |> 
  mutate(Dt = make_date(year = Year, month = Month, day = Day)) |> 
  relocate(Dt, .after = Day)

## View(arvedsn_dt_tbl)
head(arvedsn_dt_tbl)
```

\

## Create a wide version of the table

Guide to pivoting: <https://tidyr.tidyverse.org/articles/pivot.html>

```{r chunk04}
library(tidyr)
arvedsn_wide_tbl <- arvedsn_dt_tbl |> 
  pivot_wider(names_from = Item, values_from = Value, id_cols = c(Station, Dt))

## View(arvedsn_wide_tbl)
head(arvedsn_wide_tbl)
```

\

## Fill in Missing Evapotranspiration Values

First, we create a data frame with just Eto:

```{r chunk05}
arvedsn_et0_tbl <- arvedsn_dt_tbl |> 
  filter(Item == "DayEto") |> 
  select(Station, Dt, Item, Value, Unit)

head(arvedsn_et0_tbl)
```

\

Are there missing values?

```{r chunk06}
## View(arvedsn_et0_tbl)

summary(arvedsn_et0_tbl)
```

\

## Fix #1. Replace missing values with the mean or median

Step 1. Decide: mean or median?

```{r chunk07}
arvedsn_et0_tbl |> pull(Value) |> hist(breaks = 20)
```

\

Step 2. Compute the median

```{r chunk08}
value_median <- arvedsn_et0_tbl |> pull(Value) |> median(na.rm = TRUE)
value_median
```

\

Step 3. Substitute the median whenever `Value` is NA.

```{r chunk09}
arvedsn_et0_fix1_tbl <- arvedsn_et0_tbl |> 
  mutate(Value_Fix1 = if_else(is.na(Value), value_median, Value))

# View(arvedsn_et0_fix1_tbl) 
```

\

## Fix #2. Interpolate missing values

```{r chunk10}
library(zoo)
vals_linear_interpolation <- na.approx(arvedsn_et0_tbl$Value)
vals_linear_interpolation

# You can also do spline interpolation
# vals_spline_interpolation <- na.spline(arvedsn_et0_tbl$Value)
# vals_spline_interpolation
```
\

Add the imputed values as a new column with `bind_cols`:

```{r chunk11}
arvedsn_et0_fix2_tbl <- arvedsn_et0_fix1_tbl |> 
  bind_cols(Value_Fix2 = vals_linear_interpolation)

# View(arvedsn_et0_fix2_tbl)
head(arvedsn_et0_fix2_tbl)
```

\

This is just scratching the surface of dealing with missing values! 

See also: [imputeTS](https://steffenmoritz.github.io/imputeTS/) 

## CHALLENGE #2

There are also some missing values in the Maximum Daily Air Temperature (DayAirTmpMax). 

Decide what would be a reasonable method to deal with these, and implement it.

\

## Resampling data

Bin the Eto values into low, medium, and high:

```{r chunk12}
arvedsn_et0_fix3_tbl <- arvedsn_et0_fix2_tbl |> 
  mutate(eto_level = case_when(Value_Fix2 <= 0.14 ~ "low",
                               Value_Fix2 <= 0.19 ~ "medium",
                               TRUE ~ "high"))

# View(arvedsn_et0_fix3_tbl)
head(arvedsn_et0_fix3_tbl)
```

\

# END!

Save your work to render a HTML file



