---
title: "Homework 1 - ANSWERS"
author: "Intro to R Part 1: The Basics"
output: html_document
---

```{css echo = FALSE}
h2 {
  font-size:18px;
  font-weight:bold;
}
h4.author {
  font-weight:bold;
  color:#666;
  font-style:italic;
}
```

\

## Q1. Create a numeric vector called `x` with 1000 numbers. You can use any function or operator of your choice.

```{r}
x = runif(1000)
```

\

## Q2. Use `length()` to verify that `x` has exactly 1000 elements.

```{r}
length(x)
```

\
 
## Q3. Compute the mean, minimum, maximum and range of x.

```{r}
mean(x)
min(x)
max(x)
range(x)
```

\

## Q4. Write an expression that returns the 5th element of `x`.

```{r}
x[5]
```

\

## Q5. Write an expression that returns the 6th through 10th elements of `x`.

```{r}
x[6:10]
```

\

## Q6. Create a character vector containing the names of your top-three favorite flavors of ice cream.

```{r}
my_favorites = c("vanilla", "chocolate chip", "mocha")

my_favorites
```

\

## Q7. Write an expression that returns a vector containing the FIRST 10 UPPERCASE letters of the alphabet.

```{r}
LETTERS[1:10]
```

\

## Q8. Write an expression that returns a vector containing the LAST 5 LOWERCASE letter of the alphabet.

```{r}
letters[21:26]
```

\

## Q9. Write an expression that returns a vector containing the EVEN numbers from 1 to 20 (i.e., 2, 4, 6, 8, ..., 20).

```{r}
1:10 * 2
```

\

## Q10. Write an expression that returns a vector containing the ODD numbers from 1 to 20 (i.e., 1, 3, 5, 7, ..., 19).

```{r}
(1:10 * 2) - 1
```

\

## Q11. Generate 100 random numbers using `rnorm()`, saving the results to an object called `my_nums`.

```{r}
my_nums = rnorm(100)
```

\

## Q12 Write an expression that returns the elements of `my_nums` that are greater than 1.

```{r}
my_nums[ my_nums > 1 ]
```

## Q13 Write an expression that tells you HOW MANY elements of `my_nums` are greater than 1.

```{r}
sum(my_nums > 1)
```
## Q14. Write an expression that returns the current date.

```{r}
Sys.Date()
```

\

