---
title: "Homework 3 - ANSWERS"
author: "Intro to R Part 3: Data Wrangling"
output: html_document
---

```{css echo = FALSE}
h2 {
  font-size:18px;
  font-weight:bold;
}
h4.author {
  font-weight:bold;
  color:#666;
  font-style:italic;
}
```

## Setup

First we load some data from the San Francisco Open Data Portal. Specifically, we'll import a selection of 311 calls from San Francisco ([more info](https://data.sfgov.org/resource/vw6y-z8j6)).

```{r}
## Load packages
library(dplyr)
library(lubridate)

## Define a URL to a CSV
sf311_url <- "https://data.sfgov.org/resource/vw6y-z8j6.csv?$select=service_request_id,requested_datetime,agency_responsible,service_name,service_subtype,address,neighborhoods_sffind_boundaries,source&$limit=50000&$where=requested_datetime between '2023-10-11T00:00:00' and '2023-10-11T23:59:59'"

## Next, read in the data. 
##  - URLencode() converts spaces in the URL to numeric codes 
##  - read.csv() imports the CSV file from SF Open Data
##  - rename() changes the name of one of the columns
##  - ymd_hms() converts a character column into a date-time column
sf311_df <- read.csv(URLencode(sf311_url)) |> 
  rename(neighborhood = neighborhoods_sffind_boundaries) |> 
  mutate(requested_datetime = ymd_hms(requested_datetime))
```

\

If you'd like, un-comment the following line to view the data frame in a View pane:

```{r}
# View(sf311_df)
```

\

## Q01. How many rows and how many columns does this data frame have?

```{r}
dim(sf311_df)
```

\

## Q02. Write an expression that returns the column names of this data frame.

*Hint*: Look for a function that returns the names of a data frame

```{r}
names(sf311_df)
```

\

## Q03. Write an expression that returns the data frame sorted by the responsible agency.

```{r}
sf311_df |> arrange(agency_responsible) |> head()
```

\

## Q04. Write an expression that just returns three of columns (of your choice).

```{r}
sf311_df |> select(service_name, service_subtype, source) |> head()
```

\

## Q05. Write an expression that returns the range of the requested date & time.

```{r}
sf311_df |> pull(requested_datetime) |> range()
```

\

## Q06. Write an expression that just returns 311 calls that were submitted by phone.
 
```{r}
sf311_df |> filter(source == 'Phone') |> head()
```
 
\

## Q07. How many 311 calls were submitted by phone?

```{r}
sf311_df |> filter(source == 'Phone') |> nrow()
```

\

## Q08. Write an expression the returns complaints about barking dogs made by phone.

*Hint*: Look at the service_subtype column.

```{r}
sf311_df |> filter(source == 'Phone', service_subtype == 'noise_dog_barking')
```

\

## Q09. Write an expression that returns the number of calls per neighborhood, sorted *alphabetically by neighborhood*.

```{r}
sf311_df |> group_by(neighborhood) |> summarise(num_calls = n()) |> arrange(neighborhood)
```

\

## Q10. Write an expression that returns the number of calls per neighborhood, sorted by the *number of calls*.

```{r}
sf311_df |> group_by(neighborhood) |> summarise(num_calls = n()) |> arrange(desc(num_calls)) 
```

\

## Q11. What is the most common service requested?

```{r}
sf311_df |> group_by(service_name) |> summarize(num_recs = n()) |> arrange(desc(num_recs))
```

Or:

```{r}
sf311_df |> group_by(service_name) |> summarize(num_recs = n()) |> slice_max(num_recs, n = 1)
```

\

## Q12. What is the least common service requested?

```{r}
sf311_df |> group_by(service_name) |> summarize(num_recs = n()) |> slice_min(num_recs, n = 1)
```





